Penalized linear mixed models for structured genetic data

### 2021-03-19 main.tex

To quantify selection accuracy, we calculated the precision according to $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but the direction of effect was estimated correctly (in our example, all causal $\beta$ values were positive). \anna{Figure~\ref{fig:gw_prc}} displays each method's precision ($1 -$ the false discovery rate) for the corresponding number of selected SNPs. PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly over the first $\approx$ 100 SNPs allowed into the model. For the most part, PC-lasso outperforms LMM-lasso in this setting as well. None of the methods are able to accurately select more than a few causal SNPs with this sample size.  However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first 38 SNPs selected were all false positives.

Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}, and is presented in \anna{Figure~\ref{fig:gw_mse}}. In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso. This bias is most dramatic for the first $\approx$ 100 SNPs selected by each method. For any number of selected SNPs, however, both PC-lasso and LMM-lasso produce much more accurate estimates than the ordinary lasso.

\anna{Make sure this gets an overall figure number}
\pb{I still don't see an overall figure number?}

\begin{figure}[H]
     \centering
     \begin{subfigure}[b]{0.4\textwidth}
         \centering
         \includegraphics[width=\textwidth]{figures/figure_05a.pdf}
         \caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
        \label{fig:gw_prc}
     \end{subfigure}
    \hfill
     \begin{subfigure}[b]{0.4\textwidth}
         \centering
         \includegraphics[width=\textwidth]{figures/figure_05b.pdf}
         \caption{MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
        \label{fig:gw_mse}
     \end{subfigure}
     \begin{subfigure}[b]{0.15\textwidth}
         \centering
         \includegraphics[width=\textwidth]{figures/figure_05c.pdf}
     \end{subfigure}
\end{figure}

>>>>

To quantify selection accuracy, we calculated the precision according to $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but the direction of effect was estimated correctly (in our example, all causal $\beta$ values were positive). Figure~\ref{fig:gw_prc_mse}(a) displays each method's precision ($1 -$ the false discovery rate) for the corresponding number of selected SNPs. PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly over the first $\approx$ 100 SNPs allowed into the model. For the most part, PC-lasso outperforms LMM-lasso in this setting as well. None of the methods are able to accurately select more than a few causal SNPs with this sample size.  However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first 38 SNPs selected were all false positives.

Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}, and is presented in Figure~\ref{fig:gw_prc_mse}(b). In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso. This bias is most dramatic for the first $\approx$ 100 SNPs selected by each method. For any number of selected SNPs, however, both PC-lasso and LMM-lasso produce much more accurate estimates than the ordinary lasso.

\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/figure_05.pdf}
\caption{Precision and MSE for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_prc_mse}
\end{figure}
